---
title: "Create Example Data for AEA 2024 Reproducible Research Presentation"
author: "Steven J. Pierce"
bibliography: ../references.bib
csl: ../apa.csl
date: now
date-modified: last-modified
date-format: YYYY-MM-DD HH:mm:ss z
params:
  SourceDir: "scripts/"
  SourceFile: "Create_Example_Data.qmd"
  LogFile: "Create_Example_Data.html"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    number-depth: 3
    code-fold: show
    code-tools: true
    code-line-numbers: false
    embed-resources: true 
    anchor-sections: true
execute:
  eval: true
  echo: fenced
  output: true
  warning: true
  error: true
  include: true
knitr:
  opts_chunk: 
    message: true
---

# Purpose
This script prepares simulated, fictional example data for my AEA 2024 
presentation [@Pierce-RN8577].

# Setup

## Declare Path
This next chunk declares the path to this script relative to the project-level 
root directory. If the file is not in the right location under the project root
you'll get a warning message. This helps ensure relative paths are all working 
as expected. 

``` {r}
#| label: declare-path

# Declare path to this script relative to the project root directory.
here::i_am(path = paste0(params$SourceDir, params$SourceFile))
```

## Load Packages
Load R packages required for this script to work. 

```{r}
#| label: load-packages
library(here)           # for here(), to make code more portable 
library(devtools)       # for session_info()
library(dplyr)          # for %>%, bind_cols(), rename(), etc.
library(piercer)        # for git_report(), sim_2arm_prepost()
library(quarto)         # for quarto_version()
```

## Create Example Data
We need to create some simulated data to use as an example data frame called 
`MyStudy`. We'll assume a longitudinal randomized controlled trial design with 
2 groups x 4 sites x 2 time points, with balanced sample size of 20 
participants per group per site ($N = 160$ participants). 

Population parameters for the simulated data will assume: 

* Correlated ($r = 0.25$) pretest and posttest scores for a continuous, normally 
  distributed outcome.  
* Pretest mean of $M = 70$ at for both groups.
* Posttest means of $M = 75$ for the control group and $M = 81$ for the 
  treatment group. 
* Homogeneity of variance, with $SD = 10$ for both pretest and posttest scores
  in both groups. 
* No site effect on either pretest or posttest scores. 

```{r}
#| label: create-MyStudy 

# Example data frame 1
set.seed(642587)
MyStudy <- sim_2arm_prepost(Npg = 80, 
                            VNames = c("x", "y"),
                            R = matrix(data = c(1, 0.25, 0.25, 1), 
                                       nrow = 2, ncol = 2, byrow = TRUE),
                            MV.C = c(x = 70, y = 75),
                            MV.T = c(x = 70, y = 81),
                            SV.C = c(x = 10, y = 10),
                            SV.T = c(x = 10, y = 10)) %>% 
  bind_cols(., Site = factor(rep(c("A", "B", "C", "D"), times = 40))) %>% 
  rename(Group = Arm, 
         Score_Pre = x, 
         Score_Post = y, 
         Centered_Score_Pre = cx)
```


```{r}
#| label: save-MyStudy 

save(MyStudy, file = here("data/MyStudy.RData"))
```

# References
::: {#refs}
:::

# Software Information
Show information about the current software environment. This facilitates 
troubleshooting reproducibility issues if someone else is unable to get the same 
results I did from the same code. 

We used [R](https://www.r-project.org/) as our main computing environment and 
[Quarto](https://quarto.org/) scripts to enhance reproducibility. We used 
[RStudio](www.rstudio.org) as the editor to interface with R and Quarto. 

- Software chain:
  **qmd file > RStudio > Quarto > R > knitr > md file > Pandoc > HTML file**.
- Source file: **`r params$SourceFile`**
- Output file: **`r params$LogFile`**
- [Quarto `r quarto_version()`](https://quarto.org/) runs `*.qmd` files through 
  [R](https://www.r-project.org/) and [knitr](https://yihui.org/knitr/) to 
  produce `*.md` markdown files.
- [Pandoc `r rmarkdown::pandoc_version()`](https://pandoc.org) converts markdown 
  files (`*.md`) to other formats, including LaTeX (`*.tex`) and HTML (`*.html`) 
  among others.
  
## Versions
This document was generated using the following computational environment and 
dependencies: 

```{r}
#| label: show-version

# Get R and R package version numbers in use.
session_info()
```

## Git Details
The current Git commit details and status are:

``` {r}
#| label: git-details
#| echo: true

git_report()
```
